Edge sign prediction in general graphs and connection with Correlation Clustering

### 2017-06-27 gtx_algorithm.tex
add a note on singleton star

The \gtx{} algorithm takes as input an unweighted graph topology $G_0=(V_0, E_0)$ and produces a sequence of
graphs $\{G_t\}_{t=1}^T$ of decreasing size until each connected component of $G_0$ is reduced to a
single node. As we will prove later, after reaching this point, the algorithm has selected $|V_0| -
1$ edges that form a spanning tree of $G_0$. Intuitively, $G_{t+1}$ is obtained from $G_t$ by composing two
primitives so that we can informally write $G_{t+1} = \left(\collapseStar{} \circ
\extractStar{}\right)(G_t)$.

\extractStar{} partitions the graph $G_t$ into a set of stars and \collapseStar{} build the graph
made of those stars using the edges in $E_t$. We provide more details on those two operations in the
following, as well as their complexity analysis. Then we state formally the \gtx{} algorithm and prove
its termination and correctness.  Finally, we study its properties, such as the number of iterations
needed to finish and the stretch of the resulting tree. For simplicity and without loss of
generality, we assume that $G_0$ through $G_T$ consist of a single connected component.

@@

>>>>

The \gtx{} algorithm takes as input an unweighted graph topology $G_0=(V_0, E_0)$ and produces a sequence of
graphs $\{G_t\}_{t=1}^T$ of decreasing size until each connected component of $G_0$ is reduced to a
single node. As we will prove later, after reaching this point, the algorithm has selected $|V_0| -
1$ edges that form a spanning tree of $G_0$. Intuitively, $G_{t+1}$ is obtained from $G_t$ by composing two
primitives so that we can informally write $G_{t+1} = \left(\collapseStar{} \circ
\extractStar{}\right)(G_t)$.
% [from a CMU NIPS submission http://www.stat.cmu.edu/~arinaldo/papers/mutualfriends.pdf] We
% consider Mutual-Friends to be *an algorithmic primitive*, by which we mean a kind of subroutine for
% a more complicated function that iterates Mutual-Friends, similarly to [5].

\extractStar{} partitions the graph $G_t$ into a set of stars and \collapseStar{} build the graph
made of those stars using the edges in $E_t$. We provide more details on those two operations in the
following, as well as their complexity analysis. Then we state formally the \gtx{} algorithm and prove
its termination and correctness.  Finally, we study its properties, such as the number of iterations
needed to finish and the stretch of the resulting tree. For simplicity and without loss of
generality, we assume that $G_0$ through $G_T$ consist of a single connected component.
\marginpars{Actually, we only need to assume that $G_0$ is connected because then we can prove that
the other subsequent $G_t$ will be as well.}

@@

While we assume that $G_t$ is connected, it might happen during the execution of \extractStar{} that%
\begin{marginfigure}
  \centering
  \includegraphics[width=0.95\linewidth]{assets/tikz/gtx_singleton_tikz.pdf}
  \caption{The formation of a singleton star}
  \label{fig:gtx_singleton}
\end{marginfigure}
the degree of a node $u$ drops to zero because because all of its neighbors were claimed by the
periphery of previous stars. Such a node then forms a \emph{singleton star}, as illustrated in
\autoref{fig:gtx_singleton}, where after the creation of \starone{1} and \starone{2}, $c_3$ is the
single node of the third star \starone{3}.%
\begin{marginfigure}
  \centering
  \includegraphics[width=0.95\linewidth]{assets/tikz/gtx_noreduc_tikz.pdf}
  \caption{A case were too many singletons \enquote{waste} one iteration of \gtx{}}
  \label{fig:gtx_noreduc}
\end{marginfigure}
In the case no singleton are formed during one execution of \extractStar{}, the number of nodes is
reduced by at least $2$ (because each star is made of at least two nodes). On the other hand, we can
construct a graph with singletons where the factor of reduction in number of nodes and edges can be
arbitrarily close to $1$. Consider \autoref{fig:gtx_noreduc} and assume that first extract two stars
centered in $c_1$ and $c_2$.%
